# Threshold models

# PHYLIP & RPhylip
Follow the directions to install phylip and Rphylip. 
``` {r }
#+echo=FALSE
setwd("~/repos/threshold/R/")
library(geiger)
library(phytools)
library(MCMCglmmRAM)
library(coda)
library(treeplyr)
```

Let's start by simulating data under multivariate Brownian Motion with a known covariance structure.
``` {r }
# Simulate the tree
ntaxa <- 100
set.seed(1)
tree <- sim.bdtree(b=1, stop="taxa", n=ntaxa)
tree$edge.length <- tree$edge.length/max(branching.times(tree))

# Set up the covariance matrix for changes in the underlying liabilities
R <- matrix(c(  1,   0.7, -0.9,
                0.7,     1, -0.4,
                -0.9,   -0.4,   1), ncol=3, nrow=3, byrow=TRUE)
colnames(R) <- rownames(R) <- c("A", "B", "C")
# Simulate the data using 'sim.char()'. 
dat <- sim.char(tree, par=R, nsim=1, model="BM")[,,1]
colnames(dat) <- c("A", "B", "C")
pairs(dat)
```

Match the tree and to the liabilities and create discretized variables. 
``` {r }
td <- make.treedata(tree, dat)

td <- mutate(td, a = as.numeric(A > median(A)), b = as.numeric(B > median(B)), c = as.numeric(C > median(C)))
tree <- td$phy
traits <- as.data.frame(td$dat)
rownames(traits) <- td$phy$tip.label
```

Plot out the discrete traits
``` {r }
par(mfrow=c(1,1))
plot(tree, show.tip.label=FALSE)
tiplabels(pch=22, bg=traits[['a']])
tiplabels(pch=22, bg=traits[['b']], adj=c(0.53,0.5))
tiplabels(pch=22, bg=traits[['c']], adj=c(0.56,0.5))
```

# MCMCglmmRAM
``` {r }
library(MCMCglmmRAM)
traits$animal <- factor(tree$tip.label)
prior1 <- list(R=list(V=diag(2)*1e-15, fix=1), G=list(G1=list(V=diag(2), nu=0.002)))
mglmm.AC <- MCMCglmm(cbind(C, A)~trait-1, random=~us(trait):animal, rcov=~us(trait):units,
                     pedigree=tree, reduced=TRUE, data=traits, prior=prior1, 
                     pr=TRUE, pl=TRUE, family=c("gaussian", "gaussian"), thin=1)
summary(mglmm.AC)
plot(mglmm.AC)

prior2 <- list(R=list(V=diag(2)*1e-15, fix=1), G=list(G1=list(V=diag(2), nu=0.002, fix=2)))
mglmm.aC <- MCMCglmm(cbind(C, a)~trait-1, random=~us(trait):animal, rcov=~us(trait):units,
                  pedigree=tree, reduced=TRUE, data=traits, prior=prior2, 
                    pr=TRUE, pl=TRUE, family=c("gaussian", "threshold"), thin=1)
summary(mglmm.aC)
plot(mglmm.aC)

prior3 <- list(R=list(V=diag(2)*1e-15, fix=1), G=list(G1=list(V=diag(2), nu=0.002)))
mglmm.ac <- MCMCglmm(cbind(c, a)~trait-1, random=~corg(trait):animal, rcov=~corg(trait):units,
                     pedigree=tree, reduced=TRUE, data=traits, prior=prior3, 
                     pr=TRUE, pl=TRUE, family=c("threshold", "threshold"), thin=1)
summary(mglmm.ac)
plot(mglmm.ac)


prior3 <- list(R=list(V=diag(3)*1e-15, fix=1), G=list(G1=list(V=diag(3), nu=0.002)))
mglmm.abc <- MCMCglmm(cbind(a,b,c)~trait-1, random=~corg(trait):animal, rcov=~corg(trait):units,
                     pedigree=tree, reduced=TRUE, data=traits, prior=prior3, 
                     pr=TRUE, pl=TRUE, family=c("threshold", "threshold", "threshold"), thin=1)
summary(mglmm.abc)
plot(mglmm.abc)


postburnin <- floor(0.3*nrow(mglmm.ac$Liab)):nrow(mglmm.ac$Liab)
a.liab <- apply(mglmm.ac$Liab[postburnin,1:100], 2, mean)
c.liab <- apply(mglmm.ac$Liab[postburnin,101:200], 2, mean)

plot(a.liab, traits$A, pch=21,  bg=traits$a)
plot(c.liab, traits$C, pch=21,  bg=traits$c)
```

Now get the liabilities at the nodes.
``` {r }
a.nodeLiab <- apply(mglmm.ac$Sol[postburnin, 1:tree$Nnode], 2, mean)
```

And we can plot them on the tree:
``` {r }
plot(tree)
a.pie <- (a.nodeLiab-min(a.nodeLiab))/diff(range(a.nodeLiab))
a.pie <- cbind(a.pie, 1-a.pie)
nodelabels(pie=a.pie, piecol = c("white", "black"), cex=0.5)
tiplabels(pch=21, bg = c("white", "black")[traits$a+1])
```
